\documentclass{beamer}
\usepackage[utf8]{inputenc}

\usetheme{Madrid}
\usecolortheme{default}
\usepackage{amsmath,amssymb,amsfonts,amsthm}
\usepackage{txfonts}
\usepackage{tkz-euclide}
\usepackage{listings}
\usepackage{adjustbox}
\usepackage{array}
\usepackage{tabularx}
\usepackage{gvv}
\usepackage{lmodern}
\usepackage{circuitikz}
\usepackage{tikz}
\usepackage{graphicx}

\setbeamertemplate{page number in head/foot}[totalframenumber]

\usepackage{tcolorbox}
\tcbuselibrary{minted,breakable,xparse,skins}



\definecolor{bg}{gray}{0.95}
\DeclareTCBListing{mintedbox}{O{}m!O{}}{%
  breakable=true,
  listing engine=minted,
  listing only,
  minted language=#2,
  minted style=default,
  minted options={%
    linenos,
    gobble=0,
    breaklines=true,
    breakafter=,,
    fontsize=\small,
    numbersep=8pt,
    #1},
  boxsep=0pt,
  left skip=0pt,
  right skip=0pt,
  left=25pt,
  right=0pt,
  top=3pt,
  bottom=3pt,
  arc=5pt,
  leftrule=0pt,
  rightrule=0pt,
  bottomrule=2pt,
  toprule=2pt,
  colback=bg,
  colframe=orange!70,
  enhanced,
  overlay={%
    \begin{tcbclipinterior}
    \fill[orange!20!white] (frame.south west) rectangle ([xshift=20pt]frame.north west);
    \end{tcbclipinterior}},
  #3,
}
\lstset{
    language=C,
    basicstyle=\ttfamily\small,
    keywordstyle=\color{blue},
    stringstyle=\color{orange},
    commentstyle=\color{green!60!black},
    numbers=left,
    numberstyle=\tiny\color{gray},
    breaklines=true,
    showstringspaces=false,
}
%------------------------------------------------------------

\title
{12.185}
\date{October 1, 2025}
\author 
{AI25BTECH11003 - Bhavesh Gaikwad}



\begin{document}


\frame{\titlepage}
\begin{frame}{Question}
If the rank of a (5$\times$6) matrix $\vec{Q}$ is 4, then which one of the following statements is correct?

\hfill{(EE 2008)}

\begin{itemize}
    \item[a)]$\vec{Q}$ will have four linearly independent rows and four linearly independent columns.
    \item[b)]$\vec{Q}$ will have four linearly independent rows and five linearly independent columns.
    \item[c)]$\vec{Q}\vec{Q}^\top$ will be invertible.
    \item[d)]$\vec{Q}^\top\vec{Q}$ will be invertible 
\end{itemize}

\end{frame}


\begin{frame}[fragile]
    \frametitle{Theoretical Solution}
\textbf{Primary Analysis:}\\
Since rank($\vec{Q}$)=4 $\Rightarrow \; \therefore \vec{Q}$ will have four linearly independent rows and four linearly independent columns.\\

Option-A:\\
Correct Option by Primary Analysis itself.\\

Option-B:\\
Incorrect Option by Primary Analysis itself.\\

Option-C:\\
$\vec{Q}\vec{Q}^\top$ is a 5$\times$5 matrix.\\
Since, rank($\vec{Q}\vec{Q}^\top$) = rank($\vec{Q}$).\\
$\therefore$ rank($\vec{Q}\vec{Q}^\top$) = 4.\\
Since rank($\vec{Q}\vec{Q}^\top$)=4$<$5, Thus the 5$\times$5 matrix $\vec{Q}\vec{Q}^\top$ is singular $\left(|\vec{Q}\vec{Q}^\top|=0\right)$, hence not invertible.
\bigskip
Incorrect Option.\\
\end{frame}



\begin{frame}[fragile]
    \frametitle{Theoretical Solution}
Option-D:\\
$\vec{Q}^\top\vec{Q}$ is a 6$\times$6 matrix.\\
Since, rank($\vec{Q}^\top\vec{Q}$) = rank($\vec{Q}$).\\
$\therefore$ rank($\vec{Q}^\top\vec{Q}$) = 4.\\
Since rank($\vec{Q}^\top\vec{Q}$)=4$<$6, Thus the 6$\times$6 matrix $\vec{Q}^\top\vec{Q}$ is singular $\left(|\vec{Q}^\top\vec{Q}|=0\right)$, hence not invertible.
\bigskip
Incorrect Option.\\\\

Thus, Only Option-A is correct.
\end{frame}

\begin{frame}[fragile]
    \frametitle{Proof by Example}
Consider the $5\times6$ matrix $\vec{Q}$ of rank 4:
\begin{equation}
\vec{Q}
=
\myvec{
1 & 0 & 0 & 0 & 2 & 3\\
0 & 1 & 0 & 0 & 4 & 5\\
0 & 0 & 1 & 0 & 6 & 7\\
0 & 0 & 0 & 1 & 8 & 9\\
0 & 0 & 0 & 0 & 0 & 0
}.
\end{equation}

Option (a): Four independent rows and columns\\
Clearly 4 rows of $\vec{Q}$ are linearly independent. Thus row rank of $\vec{Q}$ = 4\\
Clearly 4 columns of $\vec{Q}$ are linearly independent.\\ 
Thus column rank of $\vec{Q}$ =4\\
Thus (a) holds.\\\\
Option (b): Four independent rows and Five independent columns\\
Column rank cannot exceed 4.\\
Hence (b) is false.\\\\
\end{frame}

\begin{frame}[fragile]
    \frametitle{Proof by Example}

Option (c): Invertibility of $\vec{Q}\,\vec{Q}^\top$\\

\begin{equation}
\vec{Q}\,\vec{Q}^\top
=
\myvec{
1&0&0&0&2&3\\
0&1&0&0&4&5\\
0&0&1&0&6&7\\
0&0&0&1&8&9\\
0&0&0&0&0&0
}
\,
\myvec{
1&0&0&0&0\\
0&1&0&0&0\\
0&0&1&0&0\\
0&0&0&1&0\\
2&4&6&8&0\\
3&5&7&9&0
}
\end{equation}
\begin{equation}
    \vec{Q}\vec{Q}^\top =
\myvec{
14&23&33&43&0\\
23&42&59&77&0\\
33&59&86&111&0\\
43&77&111&146&0\\
0&0&0&0&0
}
\end{equation}
Since $|\vec{Q}\,\vec{Q}^\top|=0$. Not invertible.\\ 
(c) is false.
\end{frame}

\begin{frame}[fragile]
    \frametitle{Proof by Example}
Option (d): Invertibility of $\vec{Q}^\top\vec{Q}$\\
\begin{equation}
\vec{Q}^\top\vec{Q}
=
\myvec{
1&0&0&0&0\\
0&1&0&0&0\\
0&0&1&0&0\\
0&0&0&1&0\\
2&4&6&8&0\\
3&5&7&9&0
}
\,
\myvec{
1&0&0&0&2&3\\
0&1&0&0&4&5\\
0&0&1&0&6&7\\
0&0&0&1&8&9\\
0&0&0&0&0&0
}
\end{equation}

\begin{equation}
    \vec{Q}^\top\vec{Q}=
\myvec{
1&0&0&0&2&3\\
0&1&0&0&4&5\\
0&0&1&0&6&7\\
0&0&0&1&8&9\\
2&4&6&8&120&154\\
3&5&7&9&154&197
}
\end{equation}
Since, $|\vec{Q}^\top\vec{Q}|=0$. Not invertible.\\ 
(d) is false.
\end{frame}


\end{document}